Adaptive Waveform Inversion for Transmitted Wave Data

### 2024-11-07 repro/papers/awi/paper.tex
more

AWI modifies the least-squares FWI problem by interposing a convolution operator or
{\em filter} $u \in L^2(X_{sr} \times \bR)$. That is, the residual
data $F[\kappa]-d$ is replaced by the filtered residual data $F[m]*u-d$.
Since $F[\kappa] \in L^1(X_{sr}\times \bR)$, the filtererd residual
makes sense.
The filter $u$ is intended to shift localized wave energy in $F[m]$ to
match that present in $d$. If on the other hand $u(\bx_r,t;\bx_s) =
\delta(t)$ for each $(\bx_x,\bx_r) \in X_{sr}$, then $tu = 0$. This
observation motivates the definition of the AWI objective function
\begin{equation}
  \label{eqn:awidefmain}
  J_{\rm AWI,\sigma}[\kappa;d] = \sum_{(\bx_s,\bx_r) \in X_{sr}}
  \frac{\|Tu_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2}{\|u_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2}.
\end{equation}
in which $Tu(\bx_r,t;\bx_s) = tu(\bx_r,t;\bx_s)$ and
\begin{equation}
  \label{eqn:reg}
  u_{\sigma}[\kappa;d] = \mbox{argmin}_{u \in L^2(X_{sr} \times \bR)} (\|F[\kappa]*u-d\|^2 + \sigma\|u\|^2)
\end{equation}
The objective function for the closely related MSWI approach,
mentioned in the introduction, is  
\begin{equation}
  \label{eqn:mswidefmain}
  J_{\rm MSWI,\sigma}[\kappa;d] = \sum_{(\bx_s,\bx_r) \in X_{sr}}
  \|Tu_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2.
\end{equation}
Note that while the Lax-Milgram Theorem (\cite{Yosida}, section
III.7) ensures the existence of a unique solution
$u_{\sigma}[\kappa;d] \in L^2(X_{sr}\times \bR)$ of the least squares
problem \ref{eqn:reg},
it is not immediiately obvious that $Tu_{\sigma}[\kappa;d] \in L^2(X_{sr}\times \bR)$.

@@

\begin{prop}
  \label{thm:noisefree}
  Assume Exponential Decay, Single Arrival, and No Return. Then for
  $\log \kappa \in M$,r $p \ge 1$, $F[\kappa], TF[\kappa]
  \in L^p(X_{sr} \times \bR)$, and there exist $C_0, C_1, C_2 >0$ so
  that for each $(\bx_s,\bx_r) \in X_{sr}$,
  \begin{eqnarray}
    \label{eqn:dnorms}
    C_1\|\partial_tf\|_{L^p(\bR)} - C_2\|f\|_{L^p(\bR)} \le \|F[\kappa](\bx_r,\cdot;\bx_s)\|_{L^p(\bR)} & \le & C_0\|\partial_tf\|_{L^p(\bR)}, \nonumber \\
    \|TF[\kappa](\bx_r,\cdot;\bx_s)\|_{L^p(\bR)} & \le & C_0\|\partial_tf\|_{L^p(\bR)}.
  \end{eqnarray}
\end{prop}

\begin{proof}
  A simple estimate shows that $F[\kappa]$ satisfies an exponential
  decay estimate of the form G4, whence the first estimate
  follows. Since $TbH \in W^{1,1}(\bR)$,
  \[
    T\rho(a\delta + bH)*\delta_{\tau}*\partial_t f = \partial_t TbH *
    \delta_{\tau} * f + \tau \rho (a\delta +
    bH)*\delta_{\tau}*\partial_t f  + (a\delta +
    bH)*\delta_{\tau}*T\partial_t f.
  \]
  All terms involve convolution of $f$ or $\partial_t f$ with kernels
  uniformly $L^1$-bounded in $\kappa \in M$, so the second estimate follows. 
\end{proof}

@@

\subsection{Linking Regularization and Wavelength; Statement of the
  Main Result}
In the remainder of this section, I will examine the $\lambda$-parametrized families of
objective functions obtained by
\begin{itemize}
\item[1. ] choosing the Tihonov regularization weight in the least
  square problem \ref{eqn:reg} as $\sigma = r\lambda$, in which $r$ is
  a positive constant with dimensions pressure$^2 \times $ time$^2$.
\item[2. ]replacing $S$, $d$ with $\Sl$, $\dl =\Fl[\kappa^*]$
  (noise-free data) in the definition of $\ul$ (equation \ref{eqn:reg}).
\end{itemize}

Justification for choosing the regularization weight $\sigma$ to be
proportional to
the wavelength proxy $\lambda$ will emerge from the analysis relating
$\|T\ul[\kappa;\dl]\|$ to the travel
time error $|\tau[\kappa]-\tau[\kappa^*]$. The necessary units for the
proportionality constant $r$ follow from the discussion after equation
\ref{eqn:reg}. It will be clear that choosing $\sigma$ a function of
$\lambda$ roughly proportional to $\lambda$, that is, $C_*\lambda \le
\sigma(\lambda) \le C^*\lambda$ for $C_*,C^*$ satisfying $0 < C_* \le
C^*$, would lead to the same asymptotic estimates for the relation
between the adaptive filter $\ul$ and travel time
error. The argument is slightly cleaner if $\sigma$ is simply
proportional to (non-dimensional) $\lambda$ by a necessarily
dimensional constant..

@@

>>>>

AWI modifies the least-squares FWI problem by interposing a convolution operator or
{\em filter} $u \in L^2(X_{sr} \times \bR)$. That is, the residual
data $F[\kappa]-d$ is replaced by the filtered residual data $F[\kappa]*u-d$.
Since $F[\kappa] \in L^1(X_{sr}\times \bR)$, the filtererd residual
makes sense.
The filter $u$ is intended to shift localized wave energy in $F[m]$ to
match that present in $d$. If on the other hand $u(\bx_r,t;\bx_s) =
\delta(t)$ for each $(\bx_x,\bx_r) \in X_{sr}$, then $tu = 0$. This
observation motivates the definition of the AWI objective function
\begin{equation}
  \label{eqn:awidefmain}
  J_{\rm AWI,\sigma}[\kappa;d] = \sum_{(\bx_s,\bx_r) \in X_{sr}}
  \frac{\|Tu_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2}{\|u_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2}.
\end{equation}
in which $Tu(\bx_r,t;\bx_s) = tu(\bx_r,t;\bx_s)$ and
\begin{equation}
  \label{eqn:reg}
  u_{\sigma}[\kappa;d] = \mbox{argmin}_{u \in L^2(X_{sr} \times \bR)} (\|F[\kappa]*u-d\|^2 + \sigma\|u\|^2)
\end{equation}
The objective function for the closely related MSWI approach,
mentioned in the introduction, is  
\begin{equation}
  \label{eqn:mswidefmain}
  J_{\rm MSWI,\sigma}[\kappa;d] = \sum_{(\bx_s,\bx_r) \in X_{sr}}
  \|Tu_{\sigma}[\kappa;d](\bx_r,\cdot;\bx_s)\|^2.
\end{equation}
Note that while the Lax-Milgram Theorem (\cite{Yosida}, section
III.7) ensures the existence of a unique solution
$u_{\sigma}[\kappa;d] \in L^2(X_{sr}\times \bR)$ of the least squares
problem \ref{eqn:reg},
it is not immediiately obvious that $Tu_{\sigma}[\kappa;d] \in L^2(X_{sr}\times \bR)$.

@@

\begin{prop}
  \label{thm:noisefree}
  Assume Exponential Decay, Single Arrival, and No Return. Then for
  $\log \kappa \in M$, $p \ge 1$, $F[\kappa], TF[\kappa]
  \in L^p(X_{sr} \times \bR)$, and there exist $C_0, C_1, C_2 >0$ so
  that for each $(\bx_s,\bx_r) \in X_{sr}$,
  \begin{eqnarray}
    \label{eqn:dnorms}
    C_1\|\partial_tf\|_{L^p(\bR)} - C_2\|f\|_{L^p(\bR)} \le \|F[\kappa](\bx_r,\cdot;\bx_s)\|_{L^p(\bR)} & \le & C_0\|\partial_tf\|_{L^p(\bR)}, \nonumber \\
    \|TF[\kappa](\bx_r,\cdot;\bx_s)\|_{L^p(\bR)} & \le & C_0\|\partial_tf\|_{L^p(\bR)}.
  \end{eqnarray}
\end{prop}

\begin{proof}
  A simple estimate shows that $F[\kappa]$ satisfies an exponential
  decay estimate of the form \ref{eqn:best} whence the first estimate
  follows. Since $TbH \in W^{1,1}(\bR)$,
  \[
    T\rho(a\delta + bH)*\delta_{\tau}*\partial_t f = \partial_t TbH *
    \delta_{\tau} * f + \tau \rho (a\delta +
    bH)*\delta_{\tau}*\partial_t f  + (a\delta +
    bH)*\delta_{\tau}*T\partial_t f.
  \]
  All terms involve convolution of $f$ or $\partial_t f$ with kernels
  uniformly $L^1$-bounded in $\kappa \in M$, so the second estimate follows. 
\end{proof}

@@

\subsection{Linking Regularization and Wavelength}
Note that noise-free data $\dl = \Fl[\kappa]$ is bounded in $L^2(X_{sr}
\times \bR)$ uniformly in $\kappa$ and $\lambda$  (Proposition
\ref{eqn:noisefree}, estimates \ref{eqn:Frembd}). 
The minimum value of the right-hand side in equation \ref{eqn:regs} is
the length of the projection of $(\dl,0)^T$ onto the range of
$(\Sl[\kappa[],\sigma I)^T$, so is less than $\|\dl^2\|$.

Suppose conversely that $\Sl[\kappa]\ul[\kappa;d]\| \ge
\frac{1}{2}\|d\|$. Then
\[
  \|d\| \le 2\|\Fl[\kappa]*\ul[\kappa;d] \le
  2\lambda^{\frac{1}{2}}\|\ul[\kappa;d]\|
\]
from Young's inequality, and from the bound \ref{eqn:unorm} this is in
turn
\[
  \le C \frac{\lambda}{\sigma} \|d\|
\]
whence
\[
  \sigma \le C \lambda.
\]
This inequality suggests that in order to expect filtered predicted data
$\Sl[\kappa]\ul[\kappa;d\|$ to remian uniformly (in $\lambda$) close to the data to be
inverted, as the heuristic justification for AWI suggests, the
regularization weight $\sigma$ must be coupled to the wavelength proxy
$\lambda$.  It will be clear from the discussion to follow that choosing $\sigma$ a function of
$\lambda$ roughly proportional to $\lambda$, that is, $C_*\lambda \le
\sigma(\lambda) \le C^*\lambda$ for $C_*,C^*$ satisfying $0 < C_* \le
C^*$, would suffice. The argument is slightly cleaner if $\sigma$ is simply
proportional to (non-dimensional) $\lambda$ by a necessarily
dimensional constant.. Therefore we assume from now on that
\begin{equation}
  \label{eqn:lamsig}
  \sigma = r \lambda
\end{equation}
in which $r$ is a positive constant with dimensions pressure$^2 \times
$ time$^2$.

This choice will eventually be justified by a converse estimate, that
is, showing that under the condition \ref{eqn:lamsig}, the filtered
predicted data $\Sl[\kappa]\ul[\kappa;d]$ has norm at least half that
of $\|\dl\|$ for all $\lambda \in (0,1]$, for properly chosen $r$.

@@

\section{BLAH BLAH}

Assume as in the last section, that is, that solutions
of the acoustic system \ref{eqn:awe} is well-approximated by the
geometric asymptotics as in equation \ref{eqn:hadamard}, and that the
data is noise-free ($d=F[\kappa^*]$). Together with the properties of $w$
just listed, these assumptions imply that
\begin{equation}
  \label{eqn:eq11}
  J_{0,\sigma}[\kappa;d] = O(\sigma) + O(\lambda(w)).
\end{equation}

\begin{proof}
  Since the arguments of $\Sl$,... are just along for the ride
  suppress them, as before.
  First analyse the leading order approximation: from equations
  \ref{eqn:hatuf} and \ref{eqn:hatgf},
  \[
  \|\Sla \uzl -\dzl\|^2 = \frac{(a^*)^2}{2\pi} \int d\omega \left| \fwl(\omega) \frac{|\fwl(\omega)|^2}{|\fwl(\omega)|^2 + \frac{r\lambda}{a^2}}
    - \fwl(\omega)\right|^2
\]
\[
  =r^2\lambda^2\left(\frac{(a^*)^2}{2\pi a^4}\right) \int d\omega \frac{| \fwl(\omega)|^2}
  {(|\fwl(\omega)|^2 + \frac{r\lambda}{a^2})^2}
\]
\[
  =r^2\lambda^2\left(\frac{(a^*)^2}{2\pi a^4}\right) \int d\omega
  \frac{| \lambda^{\frac{1}{2}}w_1(\lambda \omega)|^2}
  {(|\lambda^{\frac{1}{2}}w_1(\lambda \omega)|^2 + \frac{r\lambda}{a^2})^2}
\]
\[
  = \frac{1}{2\pi}\left(\frac{ra^*}{a}\right)^2 \int d\omega
  \frac{|aw_1(\omega)|^2}
  {(|aw_1(\omega)|^2 + r)^2}
\]
which depends on $M$ (and $r$) but not on $\lambda$. From the result
\ref{eqn:u0snormlam}, the same is true of the second term in the
objective function defined in the third equation  of the display
\ref{eqn:leading}: that is,
\begin{equation}
  \label{eqn:leadingconst}
  \|\Sla \uzl -\dzl\|^2 + r\lambda \|\uzl\|^2 = C
\end{equation}

\end{proof}
To see this, begin with the definition:
\[
  \tilde{J}_{0,\sigma}[\kappa;d] = \frac{1}{2}
  (\|S_{\lambda}[\kappa]u_0[\kappa_{\lambda};d]-d\|^2 + \lambda \|u_0[\kappa_{\lambda};d]\|^2.
\]
Using the approximate identities \ref{eqn:hatuf} and \ref{eqn:hatgf} and
rearranging,
\[
\|\Sl[\kappa]u_0[\kappa_;d]-d\|^2 
  \approx \int d\omega
  |a\hat{w}\hat{u_0[\kappa_{\lambda};d]}e^{i\omega\tau}-a^*\hat{w}e^{i\omega\tau^*}|^2
\]
\[
  = \int d\omega
  (a^*)^2|\hat{w}|^2\left(\frac{\lambda}{a^2|\hat{w}|^2 +
      \lambda}\right)^2   
\le \int d\omega (a^*)^2|\hat{w}|^2
  \left(\frac{\lambda}{a^2|\hat{w}|^2 + \lambda}\right)
\]
\begin{equation}
  \label{eqn:eq12}
  = \lim_{\Omega \rightarrow \infty} \int_{-\Omega}^{\Omega} d\omega \left(\frac{a^*}{a}\right)^2\lambda
  \left(\frac{a^2|\hat{w}|^2}{a^2|\hat{w}|^2 + \lambda}\right)
\end{equation}
For any $\Omega > 0$,
\[
\int_{-\Omega}^{\Omega} d\omega \left(\frac{a^*}{a}\right)^2\lambda\left(\frac{a^2|\hat{w}|^2}{a^2|\hat{w}|^2 + \lambda}\right)
\le \int_{-\Omega}^{\Omega} d\omega \left(\frac{a^*}{a}\right)^2 \lambda
  \left(\frac{ a|\hat{w}|}{(a^2|\hat{w}|^2 +
      \lambda)^{\frac{1}{2}}}\right)
\]
\[
  = \sigma \int_{-\Omega}^{\Omega} d\omega \left(\frac{a^*}{a}\right)^2 a|\hat{w}|
  \left(\frac{\sigma }{(a^2|\hat{w}|^2 +\lambda)^{\frac{1}{2}}}\right)
  \le \sigma \int_{-\Omega}^{\Omega} d\omega \frac{(a^*)^2}{a}
  |\hat{w}|
  \le \sigma \frac{(a^*)^2}{a} \|\hat{w}\|_{L^1(\bR)}.
\]
Together with the approximate inequality \ref{eqn:eq12}, this
inequality implies
that
\[
  \|S_{\lambda}[\kappa]u_0[\kappa_{\lambda};d] - d\|^2 = O(\sigma) + O(\lambda(w)).
\]
A similar argument leads to the conclusion that
\[
  \lambda\|u_0[\kappa_{\lambda};d]\|^2 = O(\sigma) + O(\lambda(w)).
\]
Together, these conclusions imply the assertion \ref{eqn:eq11}.
